### 2022-03-04 manuscript/preamble.tex
adding references

\DeclareUnicodeCharacter{0301}{*************************************}