### 2022-05-31 manuscript/table-fringillidae-small-example.tex
creating table 1 as .tex in small example vignette, and adding it to manuscript as graphic

\begin{table}

\caption{\label{tab:table1}Table 1. Ages of congruified nodes.}
\centering
\begin{tabular}[t]{l|c|c|c|c|c}
\hline
  & Node Name & taxon A & taxon B & Node Age & Study chronogram\\
\hline
1 & n1 & Emberiza citrinella & Platyspiza crassirostris & 9.6509881 & Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, A. O. Mooers. 2012. The global diversity of birds in space and time. Nature 491 (7424): 444-448\\
\hline
2 & n1 & Pheucticus tibialis & Schoeniclus elegans & 14.3336520 & Barker, F. Keith, Kevin J. Burns, John Klicka, Scott M. Lanyon, Irby J. Lovette. 2015. New insights into New World biogeography: An integrated view from the phylogeny of blackbirds, cardinals, sparrows, tanagers, warblers, and allies. The Auk 132 (2): 333-348.\\
\hline
3 & n1 & Rhodothraupis celaeno & Schoeniclus elegans & 16.2984859 & Barker, F. Keith, Kevin J. Burns, John Klicka, Scott M. Lanyon, Irby J. Lovette. 2015. New insights into New World biogeography: An integrated view from the phylogeny of blackbirds, cardinals, sparrows, tanagers, warblers, and allies. The Auk 132 (2): 333-348.\\
\hline
4 & n1 & Emberiza citrinella & Platyspiza crassirostris & 16.9499615 & Hooper, Daniel M., Trevor D. Price. 2017. Chromosomal inversion differences correlate with range overlap in passerine birds. Nature Ecology \& Evolution 1 (10): 1526-1534\\
\hline
5 & n1 & Rhodothraupis celaeno & Schoeniclus elegans & 21.5140867 & Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, A. O. Mooers. 2012. The global diversity of birds in space and time. Nature 491 (7424): 444-448\\
\hline
6 & n1 & Emberiza leucocephalos & Platyspiza crassirostris & 22.0000000 & Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845\\
\hline
7 & n1 & Emberiza citrinella & Platyspiza crassirostris & 22.3757277 & Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845\\
\hline
8 & n2 & Platyspiza crassirostris & Rhodothraupis celaeno & 7.9691925 & Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, A. O. Mooers. 2012. The global diversity of birds in space and time. Nature 491 (7424): 444-448\\
\hline
9 & n2 & Platyspiza crassirostris & Rhodothraupis celaeno & 19.7085900 & Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845\\
\hline
10 & n2 & Pheucticus tibialis & Platyspiza crassirostris & 44.2958603 & Burns, Kevin J., Allison J. Shultz, Pascal O. Title, Nicholas A. Mason, F. Keith Barker, John Klicka, Scott M. Lanyon, Irby J. Lovette. 2014. Phylogenetics and diversification of tanagers (Passeriformes: Thraupidae), the largest radiation of Neotropical songbirds. Molecular Phylogenetics and Evolution 75: 41-77.\\
\hline
11 & n3 & Pheucticus tibialis & Rhodothraupis celaeno & 10.5304440 & Barker, F. Keith, Kevin J. Burns, John Klicka, Scott M. Lanyon, Irby J. Lovette. 2015. New insights into New World biogeography: An integrated view from the phylogeny of blackbirds, cardinals, sparrows, tanagers, warblers, and allies. The Auk 132 (2): 333-348.\\
\hline
12 & n3 & Pheucticus tibialis & Rhodothraupis celaeno & 10.5379092 & Barker, F. K., K. J. Burns, J. Klicka, S. M. Lanyon, I. J. Lovette. 2013. Going to extremes: contrasting rates of diversification in a recent radiation of New World passerine birds. Systematic Biology 62 (2): 298-320.\\
\hline
13 & n3 & Pheucticus tibialis & Rhodothraupis celaeno & 18.1570685 & Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845\\
\hline
14 & n4 & Emberiza citrinella & Schoeniclus elegans & 5.3238969 & Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, A. O. Mooers. 2012. The global diversity of birds in space and time. Nature 491 (7424): 444-448\\
\hline
15 & n4 & Emberiza leucocephalos & Schoeniclus elegans & 9.8622460 & Barker, F. Keith, Kevin J. Burns, John Klicka, Scott M. Lanyon, Irby J. Lovette. 2015. New insights into New World biogeography: An integrated view from the phylogeny of blackbirds, cardinals, sparrows, tanagers, warblers, and allies. The Auk 132 (2): 333-348.\\
\hline
16 & n4 & Emberiza leucocephalos & Schoeniclus elegans & 10.3391445 & Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, A. O. Mooers. 2012. The global diversity of birds in space and time. Nature 491 (7424): 444-448\\
\hline
17 & n4 & Emberiza citrinella & Schoeniclus elegans & 12.5133870 & Hooper, Daniel M., Trevor D. Price. 2017. Chromosomal inversion differences correlate with range overlap in passerine birds. Nature Ecology \& Evolution 1 (10): 1526-1534\\
\hline
18 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.1407015 & Jetz, W., G. H. Thomas, J. B. Joy, K. Hartmann, A. O. Mooers. 2012. The global diversity of birds in space and time. Nature 491 (7424): 444-448\\
\hline
19 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.1516230 & Hedges, S. Blair, Julie Marin, Michael Suleski, Madeline Paymer, Sudhir Kumar. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution 32 (4): 835-845\\
\hline
20 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.2011990 & Barker, F. Keith, Kevin J. Burns, John Klicka, Scott M. Lanyon, Irby J. Lovette. 2015. New insights into New World biogeography: An integrated view from the phylogeny of blackbirds, cardinals, sparrows, tanagers, warblers, and allies. The Auk 132 (2): 333-348.\\
\hline
21 & n5 & Emberiza citrinella & Emberiza leucocephalos & 2.2898230 & Hooper, Daniel M., Trevor D. Price. 2017. Chromosomal inversion differences correlate with range overlap in passerine birds. Nature Ecology \& Evolution 1 (10): 1526-1534\\
\hline
\end{tabular}
\end{table}