### 2022-05-31 manuscript/table-fringillidae-small-example.tex
fixing table 1 format and other edits

\begin{table}

\caption{\label{tab:table1}Ages of congruified nodes.}
\centering
\begin{tabular}[t]{l|c|c|c|c|c}
\hline
  & Node Name & taxon A & taxon B & Node Age & Study chronogram\\
\hline
1 & n1 & Emberiza citrinella & Platyspiza crassirostris & 9.6509881 & Jetz et al. 2012 - chronogram 1\\
\hline
2 & n1 & Pheucticus tibialis & Schoeniclus elegans & 14.3336520 & Barker et al. 2015 - chronogram 1\\
\hline
3 & n1 & Rhodothraupis celaeno & Schoeniclus elegans & 16.2984859 & Barker et al. 2015 - chronogram 2\\
\hline
4 & n1 & Emberiza citrinella & Platyspiza crassirostris & 16.9499615 & Hooper et al. 2017\\
\hline
5 & n1 & Rhodothraupis celaeno & Schoeniclus elegans & 21.5140867 & Jetz et al. 2012 - chronogram 2\\
\hline
6 & n1 & Emberiza leucocephalos & Platyspiza crassirostris & 22.0000000 & Hedges et al. 2015 - chronogram 2\\
\hline
7 & n1 & Emberiza citrinella & Platyspiza crassirostris & 22.3757277 & Hedges et al. 2015 - chronogram 1\\
\hline
8 & n2 & Platyspiza crassirostris & Rhodothraupis celaeno & 7.9691925 & Jetz et al. 2012 - chronogram 1\\
\hline
9 & n2 & Platyspiza crassirostris & Rhodothraupis celaeno & 19.7085830 & Jetz et al. 2012 - chronogram 2\\
\hline
10 & n2 & Platyspiza crassirostris & Rhodothraupis celaeno & 19.7085900 & Hedges et al. 2015 - chronogram 2\\
\hline
11 & n2 & Platyspiza crassirostris & Rhodothraupis celaeno & 19.7128363 & Hedges et al. 2015 - chronogram 1\\
\hline
12 & n2 & Pheucticus tibialis & Platyspiza crassirostris & 44.2958603 & Burns et al. 2014\\
\hline
13 & n3 & Pheucticus tibialis & Rhodothraupis celaeno & 10.5304440 & Barker et al. 2015 - chronogram 1\\
\hline
14 & n3 & Pheucticus tibialis & Rhodothraupis celaeno & 10.5379092 & Barker et al. 2013\\
\hline
15 & n3 & Pheucticus tibialis & Rhodothraupis celaeno & 11.2095375 & Barker et al. 2015 - chronogram 2\\
\hline
16 & n3 & Pheucticus tibialis & Rhodothraupis celaeno & 18.1570685 & Hedges et al. 2015 - chronogram 1\\
\hline
17 & n4 & Emberiza citrinella & Schoeniclus elegans & 5.3238969 & Jetz et al. 2012 - chronogram 1\\
\hline
18 & n4 & Emberiza leucocephalos & Schoeniclus elegans & 9.8622460 & Barker et al. 2015 - chronogram 1\\
\hline
19 & n4 & Emberiza leucocephalos & Schoeniclus elegans & 10.3391445 & Jetz et al. 2012 - chronogram 2\\
\hline
20 & n4 & Emberiza leucocephalos & Schoeniclus elegans & 11.7317630 & Barker et al. 2015 - chronogram 2\\
\hline
21 & n4 & Emberiza citrinella & Schoeniclus elegans & 12.5133870 & Hooper et al. 2017\\
\hline
22 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.1407015 & Jetz et al. 2012 - chronogram 1\\
\hline
23 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.1516230 & Hedges et al. 2015 - chronogram 2\\
\hline
24 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.2011990 & Barker et al. 2015 - chronogram 1\\
\hline
25 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.2409300 & Barker et al. 2015 - chronogram 2\\
\hline
26 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.2732460 & Jetz et al. 2012 - chronogram 2\\
\hline
27 & n5 & Emberiza citrinella & Emberiza leucocephalos & 0.5760260 & Hedges et al. 2015 - chronogram 1\\
\hline
28 & n5 & Emberiza citrinella & Emberiza leucocephalos & 2.2898230 & Hooper et al. 2017\\
\hline
\end{tabular}
\end{table}