### 2022-06-06 manuscript/supplementary_figured.Rmd
submission

---
title             : "Supplementary Figures \\linebreak DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life"
shorttitle        : "DateLife Supplementary Figures"

author:
  - name          : "Luna L. Sánchez Reyes"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : ""
    email         : "sanchez.reyes.luna@gmail.com"
  - name          : "Emily Jane McTavish"
    affiliation   : "1"
  - name          : "Brian O'Meara"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Science and Engineering Building 1, School of Natural Sciences, University of California, Merced"
  - id            : "2"
    institution   : "Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville"

floatsintext      : no
figurelist        : NO
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            :
  papaja::apa6_pdf:
    includes      :
      in_header: "preamble.tex"
      after_body  : ["figures-supplementary.tex"]
---

```{r setup, include = FALSE}
library("papaja")
r_refs("paper_references.bib")
```