### 2022-06-21 manuscript/fig-benchmark.Rmd
adding option to embed figures in ms

---
title: "fig2"
output: html_document
date: '2022-06-21'
---

\begin{center}
\includegraphics[width= 6.5in]{../figures/fig_runtime_main.pdf}
\end{center}
\captionof{figure}{Computation time used to process a datelife query and a datelife search across \texttt{datelife}'s chronogram database, relative to number of input taxon names. For each N = {10, 100, 200, ..., 1 000, ... , 9 000, 10 000}, we sampled N species names from the class Aves a hundred times, and then performed a datelife search with input processing, without implementing Taxon Names Resolution Service (TNRS; dark gray), and implementing TNRS (light gray). We also performed a search using a query that had been pre-processed with TNRS, for comparison (light blue).}
\label{fig:runtime_main}