### 2022-06-21 manuscript/fig1.Rmd

---
title: "fig1"
output: html_document
date: '2022-06-21'
---
\begin{center}
\includegraphics[width = 6.5in]{../figures/figure1/figure1-new.pdf}
\captionof{figure}{Stylized DateLife workflow. This shows the general workflows and analyses that can be performed with \texttt{datelife}, via the R package or through the website at \url{http://www.datelife.org/}. Details on the functions involved on each workflow are shown in \texttt{datelife}'s R package vignette.}
\label{fig:workflow}
\end{center}