### 2022-06-21 manuscript/preamble.tex

\DeclareUnicodeCharacter{0301}{*************************************}

% leave the figure captions outside\ caption{} because I want them
% to be formatted in the same way as the general text (double spaced and linenumbered)

\captionsetup[figure]{labelfont={sc},labelformat={default},labelsep=period,name={Figure}}
\usepackage{caption}
\usepackage{float}