### 2022-06-22 manuscript/fig-fringillidae-full-example.Rmd

---
title: "Figure Fringillidae full Example"
output: html_document
date: '2022-06-21'
---

\begin{minipage}{18cm}
\begin{center}
\includegraphics{../figures/figure-fringillidae/median_and_calibration_ages-congruified.png}
\end{center}

\captionof{figure}{Fringillidae median summary chronogram generated with DateLife. It has 256 tips and 233 nodes, from which 212 have age data from at least one published chronogram.}
\label{fig:fringillidages}
\end{minipage}