### 2022-06-22 manuscript/fig-workflow.Rmd

---
title: "fig1"
output: html_document
date: '2022-06-21'
---
\begin{minipage}[t]{17.3cm}
\begin{center}
\includegraphics{../figures/figure1/figure1-new.pdf}
\captionof{figure}{Main DateLife workflow. Analyses can be performed via DateLife's interactive website at \url{www.datelife.org}, or using the \texttt{datelife} R package. Details on the R functions used to perform the analyses are available from \texttt{datelife}'s R package vignettes at \url{phylotastic.org/datelife}.}
\label{fig:workflow}
\end{center}
\end{minipage}