### 2022-06-22 manuscript/preamble.tex

\DeclareUnicodeCharacter{0301}{*************************************}

% leave the figure captions outside\ caption{} if you want them
% to be formatted in the same way as the general text (double spaced and linenumbered)
% line numbers for captions trick from https://latex.org/forum/viewtopic.php?t=34614

\usepackage{ragged2e}

\usepackage{caption}

\DeclareCaptionFont{linenumbers}{\internallinenumbers}

\usepackage{float}
\restylefloat{table}

\captionsetup[figure]{font+=linenumbers, labelfont={sc}, labelformat={default}, labelsep=period, name={Figure}}

\captionsetup[table]{font+=linenumbers, labelfont={sc}, labelformat={default}, labelsep=period, name={Table}}