### 2022-06-22 manuscript/fig-benchmark.Rmd
various edits; removing tex file of figures that is no longer necessary

---
title: "fig2"
output: html_document
date: '2022-06-21'
---

\includegraphics{../figures/fig_runtime_main.pdf}
\captionof{figure}{\textit{DateLife's benchmarking results. Computation time used to process a query and a search across \texttt{datelife}'s chronogram database, relative to number of input taxon names. For each N = \{10, 100, 200, ..., 1 000, ... , 9 000, 10 000\}, we sampled N species names from the class Aves a hundred times, and then performed a datelife search processing the input names with Taxon Names Resolution Service (TNRS; light gray), and without processing names (dark gray). For comparison, we performed a search using an input that had been pre-processed with TNRS (light blue).}}
\label{fig:runtime_main}