### 2022-07-18 dryad/README.Rmd
updating dryad readme with cross validation files

---
title: 'Generating a README.txt file for the Dryad submission of data from study "DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life"'
output:
  pdf_document:
    keep_md: yes
    pandoc_args: --listings
    includes:
      in_header: preamble.tex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  results = 'hide'
)
```

# Description of README file

```{r}
line0 <- paste("This README file was generated on", Sys.Date(), "by Luna L. Sánchez Reyes, <https://orcid.org/0000-0001-7668-2528>")
```

# General Information on data set

```{r}
line1 <- "

GENERAL INFORMATION

1. Title of Dataset

    Data from: DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life.

2. Author Information

	Corresponding Researcher   
	name: Luna L. Sanchez Reyes    
  institution: University of California, Merced, USA    
  email: sanchez.reyes.luna@gmail.com

	Co-researcher 1    
	name: Emily Jane McTavish    
	institution: University of California, Merced, USA

	Co-researcher 2    
	name: Brian C. O'Meara    
	institution: University of Tennessee, Knoxville, USA

3. Date of data collection: 2022-01-28

4. Geographic location of data collection: Online

5. Funding sources that supported the collection of the data: National Science Foundation, USA

6. Recommended citation for this dataset:

    Sánchez-Reyes LL, McTavish EJ & O’Meara BC. (2019). DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life. BioRxiv, 782094.

    Data from: DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life. Dryad, Dataset

"
```

# Original file names, rename

```{r, echo = TRUE}

file.copy(from = "../tables/table-fringillidae-small-example.csv",
          to = "../dryad/Sanchez-Reyes_etal_2022_table_1.csv",
          overwrite = TRUE)
file.copy(from = "../tables/table-fringillidae-small-example-summary.csv",
          to = "../dryad/Sanchez-Reyes_etal_2022_table_2.csv",
          overwrite = TRUE)
file.copy(from = "../tables/table-fringillidae-all-congruified.csv",
          to = "../dryad/Sanchez-Reyes_etal_2022_supplementary_table_S1.csv",
          overwrite = TRUE)
file.copy(from = "../tables/table-fringillidae-all-congruified.pdf",
          to = "../dryad/Sanchez-Reyes_etal_2022_supplementary_table_S1.pdf",
          overwrite = TRUE)
file.copy(from = "../tables/table-fringillidae-all-summary.csv",
          to = "../dryad/Sanchez-Reyes_etal_2022_supplementary_table_S2.csv",
          overwrite = TRUE)
file.copy(from = ".../tables/table-fringillidae-all-summary.pdf",
          to = "../dryad/anchez-Reyes_etal_2022_supplementary_table_S2.pdf",
          overwrite = TRUE)
file.copy(from = "../figures/figure-workflow/median_chronogram.tre",
          to = "../dryad/Sanchez-Reyes_etal_2022_figure_1_chronogram_mock_example.tre",
          overwrite = TRUE)
file.copy(from = "../figures/figure-small-example/median_chronogram.tre",
          to = "../dryad/Sanchez-Reyes_etal_2022_figure_3_chronogram_small_example.tre",
          overwrite = TRUE)
file.copy(from = "../figures/figure-fringillidae/median_chronogram.tre",
          to = "../dryad/Sanchez-Reyes_etal_2022_figure_5_chronogram_finches_example.tre",
          overwrite = TRUE)
file.copy(from = "../figures/fringillidae-topologies/fringillidae_monophyletic_topology.tre",
          to = "../dryad/Sanchez-Reyes_etal_2022_figure_4A_topology_finches_mrca.tre",
          overwrite = TRUE)
file.copy(from = "../figures/fringillidae-topologies/fringillidae_paraphyletic_topology.tre",
          to = "../dryad/Sanchez-Reyes_etal_2022_figure_4B_topology_finches_ncbi.tre",
          overwrite = TRUE)
```

# Description of data files

```{r}
line2.1 <- "DATA & FILE OVERVIEW

1. Description of dataset

These data were generated to investigate and showcase the performance of the datelife R package (<https://github.com/phylotastic/datelife>).
We showcased the application of the package with one mock example and two different biological examples.

The first biological example uses datelife on a small sample of bird species.
The second one uses datelife on bird species belonging to the family Fringillidae of 'true finches', following the NCBI taxonomy.
The mock example was created by replacing species names from the small example with letters, and reducing the number of source chronograms.

We evaluated the performance of the package datelife with two analysis: a benchmarking analysis to measure computing time of functions, and a cross validation analysis to test the accuracy and precision of the functions.

2. File List:

File 1 Name: Sanchez-Reyes_etal_2022_table_1.csv    
File 1 Description: CSV (comma separated values) file with node age results from datelife's small example, presented in Table 1 of the manuscript. It contains node names, taxon names defining the nodes, the corresponding node ages (in Million years), and references of source chronograms that were congruified to the chosen tree topology to extract node ages.

File 2 Name: Sanchez-Reyes_etal_2022_table_2.csv    
File 2 Description: CSV (comma separated values) file summarizing age data (in Million years) from table 1, per node. These data are presented in Table 2 of the manuscript.

File 3 Name: Sanchez-Reyes_etal_2022_supplementary_table_S1.csv    
File 3 Description: CSV (comma separated values) file with node age results from datelife's Fringillidae example. It contains node names, taxon names defining the nodes, the corresponding node ages (in Million years), and references of source chronograms that were congruified to the chosen tree topology (shown in Figure 4B of the manuscript) to extract node ages.

File 4 Name: Sanchez-Reyes_etal_2022_supplementary_table_S1.pdf    
File 4 Description: PDF version of file Sanchez-Reyes_etal_2022_supplementary_table_S1.csv

File 5 Name: Sanchez-Reyes_etal_2022_supplementary_table_S2.csv    
File 5 Description: CSV (comma separated values) file summarizing age data (in Million years) from table S1, per node.

File 6 Name: Sanchez-Reyes_etal_2022_supplementary_table_S2.pdf    
File 6 Description: PDF file version of file Sanchez-Reyes_etal_2022_supplementary_table_S2.csv.

File 7 Name: Sanchez-Reyes_etal_2022_figure_1_chronogram_mock_example.tre    
File 7 Description: Newick version of chronogram from the mock example shown in Figure 1 of the manuscript (https://www.biorxiv.org/content/10.1101/782094v2).

File 8 Name: Sanchez-Reyes_etal_2022_figure_3_chronogram_small_example.tre    
File 8 Description: Newick version of chronogram from the small example shown in Figure 3 of the manuscript (https://www.biorxiv.org/content/10.1101/782094v2).

File 9 Name: Sanchez-Reyes_etal_2022_figure_4A_topology_finches_mrca.tre    
File 9 Description: Newick version of topology extracted from the Open Tree of Life synthetic tree, shown in Figure 4A of the manuscript (https://www.biorxiv.org/content/10.1101/782094v2).

File 10 Name: Sanchez-Reyes_etal_2022_figure_4B_topology_finches_ncbi.tre    
File 10 Description: Newick version of topology extracted from the Open Tree of Life synthetic tree, shown in Figure 4B of the manuscript (https://www.biorxiv.org/content/10.1101/782094v2).

File 11 Name: Sanchez-Reyes_etal_2022_figure_5_chronogram_finches_example.tre    
File 11 Description: Newick version of chronogram from the finch example shown in Figure 5 of the manuscript (https://www.biorxiv.org/content/10.1101/782094v2).

File 12 Name: Sanchez-Reyes_etal_2022_supplementary_figure_S1.pdf    
File 12 Description: PDf file showing results of cross validation analysis of datelife's chronogram generating functions. 
"
```

```{r}
line2.2 <- c()
# short study references:
names_phylo_all <- c("Barker et al. 2013",  # 1
                      "Barker et al. 2015 - chronogram 1",  # 2
                      "Barker et al. 2015 - chronogram 2",  # 3
                      "Burns et al. 2014",  # 4
                      "Claramunt et al. 2015",  # 5
                      "Gibb et al. 2015", # 6
                      "Hedges et al. 2015 - chronogram 1",  # 7
                      "Hedges et al. 2015 - chronogram 2",  # 8
                      "Hooper et al. 2017",  # 9
                      "Jetz et al. 2012 - chronogram 1",  # 10
                      "Jetz et al. 2012 - chronogram 2",  # 11
                      "Kimball et al. 2019 - chronogram 1", # 12
                      "Kimball et al. 2019 - chronogram 2", # 13
                      "Oliveros et al. 2019", # 14
                      "Price et al. 2014 - chronogram 1", # 15
                      "Price et al. 2014 - chronogram 2", # 16
                      "Roquet et al. 2014 - chronogram 1", # 17
                      "Roquet et al. 2014 - chronogram 2", # 18 
                      "Uyeda et al 2017")
file_index <- 12
fig_index <- 1
for (name in names_phylo_all) {
  file_index <- file_index + 1
  fig_index <- fig_index + 1
  x1 <- paste0("File ", file_index, " Name: Sanchez-Reyes_etal_2022_supplementary_figure_S", fig_index, ".jpg")
  x2 <- paste0("File ", file_index, " Description: JPG file of results of cross validation analysis of datelife chronogram generating functions, using a tree topology from ", name, ". Comparison of original chronogram (black) and the chronogram obtained using datelife (gray).\n")
  file_index <- file_index + 1
  x3 <- paste0("File ", file_index, " Name: Sanchez-Reyes_etal_2022_supplementary_figure_S", fig_index, "_cross_validated.tre")
  x4 <- paste0("File ", file_index, " Description: Newick file of chronogram cross validated using dates obtained with datelife, shown in gray on supplementary figure Sanchez-Reyes_etal_2022_supplementary_figure_S", fig_index, ".jpg\n")
  file_index <- file_index + 1
  x5 <- paste0("File ", file_index, " Name: Sanchez-Reyes_etal_2022_supplementary_figure_S", fig_index, "_source.tre")
  x6 <- paste0("File ", file_index, " Description: Newick file of source chronogram from", name, " used for the cross validaton analysis shown on supplementary figure Sanchez-Reyes_etal_2022_supplementary_figure_S", fig_index, ".jpg\n")

  line2.2 <- c(line2.2, x1, x2, x3, x4, x5, x6)
}

```

<!--
	File XXX Name: Sanchez-Reyes_etal_2022_supplementary_    
	File XXX Description:

	File XXX Name: Sanchez-Reyes_etal_2022_supplementary_    
	File XXX Description: 
-->

# Methods that generated the data files

```{r}
line3 <- "
METHODOLOGICAL INFORMATION

All age data and chronograms were obtained using the R package datelife v0.6.5 https://CRAN.R-project.org/package=datelife and the chronogram database OpenTreeChronograms v2022.1.28
"
```

# Individual descriptions of each data file

```{r}
var_description_1 <- c("Node names in tree topology, as defined by taxon A and taxon B.",
                       "Scientific name of the first element of a taxon pair that defines a node from a tree topology.",
                       "Scientific name of the second element of a taxon pair that defines a node from a tree topology.",
                       "Ages of nodes defined by taxon A and taxon B, in Million years.",
                       "Reference of the study that published the chronogram from which the node ages were extracted.")

var_description_2 <- c("Names of nodes from the tree topology used for the dating analysis.",
                       "Minimum node age, in Million years.",
                       "First quartile of node age distributions, in Million years.",
                       "Mean node age, in Million years.",
                       "Median node age, in Million years.",
                       "Third quartile of node age distributions, in Million years.",
                       "Maximum node age, in Million years.",
                       "Variance of node age, in Million years.",
                       "Standard deviation of node age, in Million years.")

csv_file_names <- c("Sanchez-Reyes_etal_2022_table_1.csv",
                    "Sanchez-Reyes_etal_2022_table_2.csv",
                    "Sanchez-Reyes_etal_2022_supplementary_table_S1.csv",
                    "Sanchez-Reyes_etal_2022_supplementary_table_S2.csv")
tre_file_names <- c("Sanchez-Reyes_etal_2022_figure_1_chronogram_mock_example.tre")

text <- c()
var_descriptions <- list(var_description_1, var_description_2, var_description_1, var_description_2) 
names(var_descriptions) <- csv_file_names

for (file_name in csv_file_names) {
  # read the csv file in:
  csv_data <- utils::read.csv(file = paste0("../dryad/", file_name), header = TRUE, row.names = NULL)
  # remove the column with row numbers:
  csv_data <- csv_data[,-1]
  # create the description text
  var_list <- gsub("[.]", " ", colnames(csv_data))
  var_list <- paste0("\t", var_list, ": ", var_descriptions[[file_name]])
  text <- c(text, 
            c(paste("\nDATA-SPECIFIC INFORMATION FOR:", file_name, "\n"),
              paste("1. Number of variables/columns:", ncol(csv_data), "\n"),
              paste("2. Number of cases/rows:", nrow(csv_data), "\n"),
              "3. Variable List:", var_list,
              "\n4. Missing data codes: None",
              "\n5. Abbreviations used: NA = not applicable\n"))
}

line4 <- text
```

# Writing everything to TXT file

```{r, echo = TRUE}

writeLines(text = c(line0, 
                    line1,
                    line2.1,
                    line2.2,
                    line3,
                    line4), con = "../dryad/README_Sanchez-Reyes_etal_2022_data.txt")
```