### 2023-01-14 manuscript/fig-workflow.Rmd
updating path to figure

---
title: "fig1"
output: html_document
date: '2022-06-21'
---
\begin{minipage}[t]{17.3cm}
\includegraphics{../figures/figure-workflow/figure1-new.pdf}
\captionof{figure}{\textit{Main DateLife workflow. Analyses can be performed via DateLife's interactive website at \url{www.datelife.org}, or using the \texttt{datelife} R package. Details on the R functions used to perform the analyses are available from \texttt{datelife}'s R package vignettes at \url{https://phylotastic.org/datelife}.}}
\label{fig:workflow}
\end{minipage}