### 2023-05-04 tables/README-tables.Rmd

---
title: "Tables"
output: html_document
date: "2023-05-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("xtable")
library(xtable)
```

Use of package `xtable` to create latex tables:
Trick from https://www.youtube.com/watch?v=N-hLp8ADhVE

```{r}
table1 <- read.csv(file = "table-fringillidae-small-example.csv")
head(table1)
print(xtable(table1, type = "latex"))
```