### 2023-06-09 tables/README-tables.Rmd
create tex file for table 2

---
title: "Tables"
output: html_document
date: "2023-05-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("xtable")
library(xtable)
```

tex file of small examples table 1 developed in vignettes/figure-small-example.Rmd L289-295.
It is the same as the tex resulting in the code at the end.

Tex file for table 2 generated here:

Use of package `xtable` to create latex tables:
Trick from https://www.youtube.com/watch?v=N-hLp8ADhVE

```{r}
table2 <- read.csv(file = "table-fringillidae-small-example-summary.csv")
head(table2)
t2 <- knitr::kable(table2, 
                   format = 'latex',
                   row.names = FALSE,
                   align = "c",
                   caption = "Summary of congruified ages per node.",
                   label = "table2")

# t2 <- xtable(table2, type = "latex")
class(t2)
writeLines(t2, '../tables/table-fringillidae-small-example-summary.tex')

```

```{r}
table1 <- read.csv(file = "table-fringillidae-small-example.csv")
head(table1)
print(xtable(table1, type = "latex"))
```