### 2023-11-15 tables/table-fringillidae-small-example-summary.Rmd
tables as doc

---
title: ""
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
table2 <- read.csv(file = "table-fringillidae-small-example-summary.csv")
head(table2)
colnames_edited <- gsub("\\.", " ", colnames(table2))
colnames_edited
```

```{r, echo =FALSE}
knitr::kable(table2, 
             digits = 3,
             row.names = FALSE,
             col.names = colnames_edited,
             align = "l",
            caption = "Table 2. Summary statistics of node ages from Table 1. DateLife uses median ages by default to date a chosen tree topology (Fig. 3g).")
```