### 2023-11-15 tables/table-fringillidae-small-example.Rmd

---
title: ""
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
table10 <- read.csv(file = "table-fringillidae-small-example.csv")
head(table10)
colnames_edited <- gsub("\\.", " ", colnames(table10))
colnames_edited 
colnames_edited[5] <- "Study reference"
```

```{r, echo =FALSE}
knitr::kable(table10, 
             digits = 3,
             row.names = FALSE,
             col.names = colnames_edited,
             align = "l",
            caption = "Table 1. Node ages extracted with DateLife from 9 chronograms (shown in Fig. 2c) containing at least 2 of the 6 bird species analysed in the Fringillidae small example. All nodes are congruent with the chosen tree topology shown in Figure 2d.")
```