### 2023-12-15 data-raw/felis-chronogram-plot.Rmd
workflow to plot calibrations for felis

---
title: "felis plot"
output: pdf_document
date: "2023-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
source(file = "R/plot_node_ages_v2.3.R")
library(ape)
library(datelife)
```

```{r}
felis_node_subtree <- rotl::tol_subtree(node_id = 'ott563165', label = "name")
my_datelife_query <- datelife::make_datelife_query(input = felis_node_subtree)
ls(my_datelife_query)
my_datelife_result <-  datelife::get_datelife_result(input = my_datelife_query, cache = opentree_chronograms)
```

```{r}
my_datelife_result$`Roquet, C., Lavergne, S., & Thuiller, W. (2014). One Tree to Link Them All: A Phylogenetic Dataset for the European Tetrapoda. PLoS Currents. doi:10.1371/currents.tol.5102670fff8aa5c918e78f5592790e48
`
```

```{r}
phylo_median <- datelife::datelife_search(input = my_datelife_query,
                                           summary_format = "phylo_median")
#dsumm <- datelife:::summary.datelifeResult(datelife_query = my_datelife_query, object = my_datelife_result)

phylo_all <- datelife::datelife_search(input = my_datelife_query,
                                           summary_format = "phylo_all")
class(phylo_all)
ls(phylo_all)
length(phylo_all)
```

### Getting oldest age among all source chronograms in a `multiPhylo` object:

```{r}
lapply(phylo_all, ape::branching.times)

plot_max_time <- 846
```

```{r}
all_calibrations <- datelife:::extract_calibrations_phylo(
  input = phylo_all,
  each = FALSE)
nrow(all_calibrations)
```
Congruify calibrations:

```{r}
congruified_calibs <- datelife::congruify_and_mrca_multiPhylo(phy = phylo_median,
                              source_chronograms = phylo_all[-9])

```

```{r}

calibrations <-  datelife:::summary.matchedCalibrations(phylo_median)
ls(calibrations)
calibrations$in_phy
congruified_calibs

calibrations <- list(in_phy = congruified_calibs)
plot_node_ages_2.3(phylo_median, calibrations)
```

```{r}
datelifeplot::plot_node_ages2(chronogram = phylo_median, matched_ages = calibrations)
```